---
title: "Demystifying R Survey Report"
output: 
    html_document:
        css: slides/style.css
        toc: true
        toc_depth: 2
        toc_float: true
        df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 3)
```

```{r}
# Load packages

library(tidyverse)
library(googlesheets)
library(janitor)
library(lubridate)
library(tigris)
library(hrbrthemes)
library(gendercodeR)
library(ggmap)
library(leaflet)
library(scales)
```

```{r}
# Set plotting default theme

theme_set(theme_ipsum(base_family = "Karla",
                      axis_title_family = "Karla"))
```


```{r}
# Download data

gs_title("Demystifying R Survey") %>%
  gs_download(to = here::here("data", "survey-responses.csv"),
              overwrite = TRUE)
```


```{r}
# Import our data

survey_responses_today <- read_csv("data/survey-responses.csv") %>% 
  # Make the variable names easy to work with
  clean_names() %>% 
  # Make sure the how_old_are_you variable is numeric
  mutate(how_old_are_you = parse_number(how_old_are_you)) %>% 
  # Convert the timestamp variable into something easier to work with
  mutate(timestamp = mdy_hms(timestamp)) %>%
  mutate(timestamp = as_date(timestamp)) %>%
  # Filter to only keep responses from today
  filter(timestamp == as_date(now()))
```

# About our Respondents

```{r}
number_of_responses <- nrow(survey_responses_today)

today_month <- month(now(),
                     label = TRUE,
                     abbr = FALSE) %>% 
  as.character()

today_day <- day(now()) %>% 
  as.character()

today_year <- year(now()) %>% 
  as.character()

today_pretty <- str_glue("{today_month} {today_day}, {today_year}")
```

We did a survey on `r today_pretty`. We received responses from `r number_of_responses` people. 




## Age

```{r}
mean_age <- survey_responses_today %>% 
  summarize(mean_age = mean(how_old_are_you, na.rm = TRUE)) %>% 
  mutate(mean_age = round_half_up(mean_age, 1)) %>% 
  pull(mean_age)

max_age <- max(survey_responses_today$how_old_are_you,
               na.rm = TRUE)

min_age <- min(survey_responses_today$how_old_are_you,
               na.rm = TRUE)
```

Respondents listed their age. The youngest respondent was `r min_age` years old and the oldest respondent was `r max_age`. The mean age was `r mean_age`.

The graph belows shows the age of respondents broken into three groups.

```{r}
age_categorized <- survey_responses_today %>% 
  mutate(age_category = case_when(
    how_old_are_you < 30 ~ "Less than 30",
    how_old_are_you < 50 ~ "Between 30 and 50",
    how_old_are_you > 50 ~ "50 or older",
  )) %>% 
  select(how_old_are_you, age_category) %>% 
  count(age_category) %>% 
  mutate(age_category = factor(age_category, levels = c("Less than 30",
                                                        "Between 30 and 50",
                                                        "50 or older"))) %>% 
  mutate(age_category = fct_rev(age_category)) %>% 
  drop_na(age_category)

ggplot(age_categorized, aes(age_category, n)) +
  geom_col(fill = "#6CABDD") +
  geom_text(aes(label = n),
            hjust = 2,
            color = "white") +
  coord_flip() +
  scale_y_continuous(breaks = pretty_breaks()) +
  labs(title = "Age of Respondents",
       x = NULL,
       y = NULL) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_blank())
```





## Education

The education levels of respondents are below.

```{r fig.height=4}
survey_responses_today %>% 
  count(what_is_the_highest_level_of_education_you_have_completed) %>% 
  mutate(what_is_the_highest_level_of_education_you_have_completed = factor(what_is_the_highest_level_of_education_you_have_completed, levels = c("Bachelor's degree", "Master's degree", "Doctoral degree", "Other"))) %>% 
  mutate(what_is_the_highest_level_of_education_you_have_completed = fct_rev(what_is_the_highest_level_of_education_you_have_completed)) %>% 
  drop_na(what_is_the_highest_level_of_education_you_have_completed) %>% 
  ggplot(aes(what_is_the_highest_level_of_education_you_have_completed, n)) +
  geom_col(fill = "#6CABDD") +
  geom_text(aes(label = n),
            hjust = 2,
            color = "white") +
  coord_flip() +
  labs(title = "Education Levels of Respondents",
       x = NULL,
       y = NULL) +
  scale_y_continuous(breaks = pretty_breaks()) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_blank())
```


## Locations

Respondents listed the location of their primary residence. The map below shows their locations. 

```{r fig.height=6}
respondent_locations <- survey_responses_today %>% 
  drop_na(what_is_the_city_state_and_country_of_your_primary_residence) %>% 
  mutate_geocode(what_is_the_city_state_and_country_of_your_primary_residence) %>% 
  select(what_is_the_city_state_and_country_of_your_primary_residence, lon, lat)

leaflet(respondent_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(~lon, ~lat,
                   color = "#FF7400",
                   stroke = FALSE, 
                   fillOpacity = 0.7,
                   popup = ~what_is_the_city_state_and_country_of_your_primary_residence,
                   clusterOptions = markerClusterOptions(showCoverageOnHover = FALSE))

```

## Gender

Respondents were asked to list their gender, in whatever form they choose to identify it. 

The original responses are below.


```{r}
survey_responses_today %>% 
  drop_na(what_is_your_gender) %>% 
  select(what_is_your_gender) %>% 
  set_names("Gender (Original Response)") 
```

The table below shows a summary of all responses.


```{r}
survey_responses_today %>% 
  drop_na(what_is_your_gender) %>% 
  select(what_is_your_gender) %>% 
  mutate(gender_recoded = recode_gender(what_is_your_gender)) %>% 
  count(what_is_your_gender) %>% 
  set_names("Gender (Original Response)", "Number of Responses") 
```

We then used the [gendercodeR package](https://github.com/ropenscilabs/gendercoder) to recode responses. The table below shows the original responses and the recoded responses.

```{r}
survey_responses_today %>% 
  drop_na(what_is_your_gender) %>% 
  select(what_is_your_gender) %>% 
  mutate(gender_recoded = recode_gender(what_is_your_gender)) %>% 
  mutate(gender_recoded = str_to_title(gender_recoded)) %>% 
  replace_na(list(gender_recoded = "Unable to recode")) %>% 
  set_names("Original Responses", "Recoded Gender") 

```

We can then summarize the recoded responses, which we do in the following table.


```{r}
survey_responses_today %>% 
  drop_na(what_is_your_gender) %>% 
  select(what_is_your_gender) %>% 
  mutate(gender_recoded = recode_gender(what_is_your_gender)) %>% 
  mutate(gender_recoded = str_to_title(gender_recoded)) %>% 
  replace_na(list(gender_recoded = "Unable to recode")) %>% 
  count(gender_recoded) %>% 
  set_names("Recoded Gender", "Number of Responses") 
```


# Familiarity with R

```{r}
mean_familiarity <- survey_responses_today %>% 
  summarize(mean_familiarity = mean(how_familiar_are_you_with_r, na.rm = TRUE)) %>% 
  mutate(mean_familiarity = round_half_up(mean_familiarity, 1)) %>% 
  pull(mean_familiarity)
```

On a 5-point scale, respondents listed their level of familiarity with R as `r number(mean_familiarity, .1)`. The figure below shows familiarity broken down by education level.

```{r fig.height = 4}
survey_responses_today %>% 
  drop_na(what_is_the_highest_level_of_education_you_have_completed) %>% 
  group_by(what_is_the_highest_level_of_education_you_have_completed) %>% 
  summarize(mean_familiarity = mean(how_familiar_are_you_with_r, na.rm = TRUE)) %>% 
  mutate(mean_familiarity = round_half_up(mean_familiarity, 1)) %>% 
  mutate(what_is_the_highest_level_of_education_you_have_completed = fct_relevel(what_is_the_highest_level_of_education_you_have_completed, levels = c("Bachelor's degree", "Master's degree", "Doctoral degree", "Other"))) %>% 
  mutate(what_is_the_highest_level_of_education_you_have_completed = fct_rev(what_is_the_highest_level_of_education_you_have_completed)) %>% 
  ggplot(aes(what_is_the_highest_level_of_education_you_have_completed, mean_familiarity)) +
  geom_col(fill = "#6CABDD") +
  geom_text(aes(label = number(mean_familiarity, .1)),
            hjust = 1.5,
            color = "white") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 5),
                     breaks = seq(0, 5, by = 1)) +
  labs(title = "Familiarity with R by Education Level",
       x = NULL,
       y = NULL) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```



# Interest in Learning R

```{r}
mean_interest <- survey_responses_today %>% 
  summarize(mean_interest = mean(how_interested_are_you_in_learning_r, na.rm = TRUE)) %>% 
  mutate(mean_interest = round_half_up(mean_interest, 1)) %>% 
  pull(mean_interest)
```

On a 5-point scale, respondents listed their level of interest in learning R as `r number(mean_interest, .1)`.

We can do a similar breakdown, but adding age. With one line of code, we can make small multiples. 

```{r fig.height = 8}
r_interest_by_ed_and_age <- survey_responses_today %>% 
  mutate(age_category = case_when(
    how_old_are_you < 30 ~ "Less than 30",
    how_old_are_you < 50 ~ "Between 30 and 50",
    how_old_are_you > 50 ~ "50 or older",
  )) %>% 
  group_by(what_is_the_highest_level_of_education_you_have_completed, age_category) %>% 
  summarize(mean_interest = mean(how_interested_are_you_in_learning_r, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(mean_interest = round_half_up(mean_interest, 1)) %>% 
  mutate(what_is_the_highest_level_of_education_you_have_completed = factor(what_is_the_highest_level_of_education_you_have_completed, levels = c("Bachelor's degree", "Master's degree", "Doctoral degree", "Other"))) %>% 
  mutate(age_category = fct_relevel(age_category, 
                                    levels = c("Less than 30",
                                               "Between 30 and 50",
                                               "50 or older"))) %>% 
  mutate(age_category = fct_rev(age_category)) %>% 
  drop_na(age_category, what_is_the_highest_level_of_education_you_have_completed)


ggplot(r_interest_by_ed_and_age, 
       aes(age_category, mean_interest,
           fill = what_is_the_highest_level_of_education_you_have_completed)) +
  geom_col() +
  geom_text(aes(label = number(mean_interest, .1)),
            hjust = 1.5,
            color = "white") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 5),
                     breaks = seq(0, 5, by = 1)) +
  scale_fill_manual(values = c("#6CABDD", "#FF7400", "#FFC659", "red")) +
  labs(title = "Interest in Learning R by\nEducation Level and Age",
       x = NULL,
       y = NULL) +
  facet_wrap(~what_is_the_highest_level_of_education_you_have_completed,
             ncol = 1) +
  theme(legend.position = "none", 
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```


